---
title: 'Peer-graded Assignment: Prediction Assignment Writeup'
author: "Ayush Mohapatra"
date: "10/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal 

The goal of your project is to predict the manner in which they did the exercise. 
Your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.


## Verification
To check the validity of the fitness devices and if they give the correct output.


## Data 
This report contains the prediction analysis and data are used from the following sites:

http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv for the training data, and 
http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv for the testing data.

The dataset consists of:

  A) The Posture is Exactly correct
  B) Throwing elbows to the front
  C) Lifting dumbbell halfway
  D) Lowering dumbell halfway
  E) Throwing hips to the front
  
  
## Downloading and Reading the data

The file is downloaded from the given websites and is stored in the variable "dataset"


```{r}
if(!file.exists("pml-training.csv"))
{
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "pml-training.csv", method = 'curl')
}
dataset <- read.csv("pml-training.csv", na.strings = c("NA", ""))
if(!file.exists("pml-testing.csv"))
{
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "pml-testing.csv", method = 'curl')
}
validation <- read.csv("pml-testing.csv")
```

## Preprocessing

Packages are installed 

```{r}
library(caret)
library(randomForest)
```


For reproducibilty, a seed is set.

```{r}
set.seed(17)
```


## Data Partitioning

```{r}
inTrain = createDataPartition(y=dataset$classe, p=0.7, list=FALSE) #70percent data goes to training set
training = dataset[inTrain,]
testing = dataset[-inTrain,]
```


Removing entries with values= NA

```{r}
naColumns = sapply(training, function(x) {sum(is.na(x))}) 
naColumns
columnsWithNA = names(naColumns[naColumns > 0]) 
training = training[, !names(training) %in% columnsWithNA] 
names(training)

training <- training[, !names(training) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

## Removing NA values for Validation and Testing Set

```{r}
naColumns = sapply(validation, function(x) {sum(is.na(x))}) 
columnsWithNA = names(naColumns[naColumns > 0]) 
validation = validation[, !names(validation) %in% columnsWithNA] 
validation <- validation[, !names(validation) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

```{r}
naColumns = sapply(testing, function(x) {sum(is.na(x))}) 
columnsWithNA = names(naColumns[naColumns > 0]) 
testing = testing[, !names(testing) %in% columnsWithNA] 
testing <- testing[, !names(testing) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

## Prediction Model

```{r acc}
model <- randomForest(classe ~ .,   data=training, ntree = 50)
predictions <- predict(model, testing)
confusionMatrix(predictions, testing$classe)
modelAcc <- confusionMatrix(predictions, testing$classe)$overall[[1]]
```

Prediction Model is accurate and the same function has to be carried out on validation set

```{r}
predictions <- predict(model, validation)
predictions
```

## Conclusion

Error in testing set is much higher than  in validation set.


